---
title: "tidy_data.Rmd"
output:
  pdf_document: default
  html_document: default
---
---
title: "tidy_data.Rmd"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
d <- readRDS("health_data.RData")
```

```{r}
d$Category   <- as.factor(d$Category)
d$StateAbbr  <- as.factor(d$StateAbbr)
d$StateDesc  <- as.factor(d$StateDesc)
d$CountyName <- as.factor(d$CountyName)
d$MeasureId  <- as.factor(d$MeasureId)
```

```{r}
d_outcome <- d[d$Category == "Health Outcomes", ]
```

```{r}
table(d_outcome$MeasureId)
```

```{r}
# Clean the Geolocation column to only have the (longitutde, latitude) coordinates
# take out the parentheses so they are only the numbers
d_outcome$Geolocation <- gsub("POINT ", "", d_outcome$Geolocation)
d_outcome$Geolocation <- gsub("\\(", "", d_outcome$Geolocation)
d_outcome$Geolocation <- gsub("\\)", "", d_outcome$Geolocation)
# str_split to split the longitude and latitude
l <- strsplit(d_outcome$Geolocation, " ", fixed = TRUE)
# Create longitude and latitude columns on the d_outcome data frame
# sapply what we get out of str_split
d_outcome$longitude <- as.numeric( sapply(l, function(x) { x[1] } ) ) # for longitude
d_outcome$latitude  <- as.numeric( sapply(l, function(x) { x[2] } ) ) # for latitude
```


```{r}
d_clean <- d_outcome %>%
  transmute('county'     = CountyName,
            'latitude'   = latitude,
            'longitude'  = longitude,
            'state_abb'  = StateAbbr,
            'state_name' = StateDesc,
            'measure_id' = MeasureId,
            'data_value' = Data_Value,
            'year'       = Year)
```


## NHGIS Tidying

```{r}
d_1 <- read.csv("nhgis0028_ds239_20185_county.csv")
```

from codebook:
AJWNE001:    Total
AJWNE002:    White alone
        AJWNE003:    Black or African American alone
        AJWNE004:    American Indian and Alaska Native alone
        AJWNE005:    Asian alone
        AJWNE006:    Native Hawaiian and Other Pacific Islander alone
        AJWNE007:    Some other race alone
        AJWNE008:    Two or more races
        
school attainment:
AJYPE001:    Total
        AJYPE002:    No schooling completed
        AJYPE003:    Nursery school
        AJYPE004:    Kindergarten
        AJYPE005:    1st grade
        AJYPE006:    2nd grade
        AJYPE007:    3rd grade
        AJYPE008:    4th grade
        AJYPE009:    5th grade
        AJYPE010:    6th grade
        AJYPE011:    7th grade
        AJYPE012:    8th grade
        AJYPE013:    9th grade
        AJYPE014:    10th grade
        AJYPE015:    11th grade
        AJYPE016:    12th grade, no diploma
        AJYPE017:    Regular high school diploma
        AJYPE018:    GED or alternative credential
        AJYPE019:    Some college, less than 1 year
        AJYPE020:    Some college, 1 or more years, no degree
        AJYPE021:    Associate's degree
        AJYPE022:    Bachelor's degree
        AJYPE023:    Master's degree
        AJYPE024:    Professional school degree
        AJYPE025:    Doctorate degree
        
median household income
AJZAE001:    Median household income in the past 12 months (in 2018 inflation-adjusted dollars)


```{r}
d_1 <- d_1 %>%
 transmute('county'     = COUNTY,
            'state_name' = STATE,
            'tot_pop'    = AJWNE001,
            'med_inc'    = AJZAE001,
            'higher_ed'  = (AJYPE017 + AJYPE018 + AJYPE019 + AJYPE020 +
           AJYPE021 + AJYPE022 + AJYPE023 + AJYPE024 + AJYPE025) / AJYPE001)
```


gsub matches each occurance of the pattern

```{r}
#setdiff(d_1$county, d_clean$county)
```

```{r}
d_1$county_name <- gsub("(\\sCounty|\\sCensus Area|\\sCity|\\sBorough|\\sMunicipality)$", "", d_1$county)
```

setdiff()

```{r}
#setdiff(d_1$county_new, d_clean$county)
```

```{r}
d_clean <- d_clean %>%
  mutate('county_name' = county)
```

```{r}
d_new <- left_join(d_clean, d_1, by = c('county_name','state_name')) %>%
  mutate(state_name = as.factor(state_name),
         county_name = as.factor(county_name),
         measure_id = as.factor(measure_id)) %>%
  select(state_name,
         county_name,
         measure_id, 
         tot_pop,
         med_inc,
         higher_ed,
         longitude,
         latitude)
```

```{r}
#d_new <- gather(d_new, key = "predictor", value = "predictor_value", 5:6)
```

```{r}
saveRDS(d_new, "d_new.RData")
```


